---
title: "Changing Temperatures in Portland, OR"
date: "Last Update: `r Sys.Date()`"
output:
    rmdformats::material:
      highlight: kate
---


```{r setup, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```
``` {r echo = FALSE, warning = FALSE, message = FALSE}
library(plotly)
library(dplyr)
library(leaflet)
library(leaflet.extras)
library(rnoaa)
library(purrr)
library(lubridate)
library(ggplot2)
library(tidyr)
library(stringr)
library(ggridges)

temp_data <- readr::read_csv('data/daily-temp-data.csv', col_names = TRUE)
#pull in the data from the csv. To update run the get-data.r script in the data folder

temp_data <- temp_data %>%
  mutate(tavg = (tmin+ tmax) / 2, year = year(date), month = month(date)) %>%
  mutate(year_month = ymd(paste(year, month, "15", sep = "-")))

station_ids <- temp_data %>%
  distinct(temp_data, station, .keep_all = TRUE) %>%
  select(-c('date', 'tmin', 'tmax'))
#select the distinct station ids from the csv. only keep the id, name, lat, long

```
# NOAA Station

Below are the stations the weather data will be pulled from. Click on the points to see more information about each station.

``` {r echo = FALSE, warning = FALSE}
station_map <- leaflet(station_ids) %>%
  setView(lng = -122.6750, lat = 45.6081, zoom = 9) %>%
  addTiles() %>%
  addMarkers(lat = ~latitude, lng = ~longitude, 
             popup = paste("Station Name: ", station_ids$name, "<br>",
                           "Elevation: ", station_ids$elevation, " ft."),
             label = station_ids$name) %>%
  suspendScroll()
#create a map centered around portland. markers are the stations. suspendScroll is working,
#but when changing sleep time, it fails
station_map
```

# Average Temperature by Month

Here are is a stacked density plot (aka a ridgeline plot) of the average temperature each month over the years. The averages displayed are the averages from the number of weather stations shown in the map.

``` {r echo = FALSE, warning = FALSE}

temp_data_melted <- gather(temp_data, key = 'measurement', value = 'temperature', 
                           c(tmin, tmax, tavg))
#pivot to long format for easier graphing. pivot along the tmin, tmax, tavg columns

avg_by_year <- temp_data_melted %>%
  group_by(year, measurement) %>%
  summarize(avg_temp = mean(temperature, na.rm = TRUE), 
            max_monthly_temp = max(temperature),
            min_monthly_temp = min(temperature)) %>%
  filter(year != 2020)

temp_data_no_2020 <- temp_data %>%
  filter(year != 2020)
```
:::: {style='display: flex;'}

::: {.column width="48%"}

``` {r echo = FALSE, warning = FALSE}

ridgeline_by_year <- ggplot(temp_data_no_2020, aes(x = tavg, y = year, group = year,
                                                   fill = stat(x))) +
  geom_density_ridges_gradient(scale = 7, rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "Temp (F)", option = "C") +
  labs(x = "Temperature (F)", y = "Year",
       title = "Greater Portland Area Temperatures")
ridgeline_by_year
```

:::

:::{.column width="4%"}
:::

::: {.column width="48%"}

``` {r echo = FALSE, warning = FALSE}
avg_by_year_plot <- ggplot(avg_by_year, aes(x = year, y = avg_temp))

avg_by_year_plot <- avg_by_year_plot +
  geom_line(aes(group = measurement, color = measurement)) +
  geom_smooth(method = "lm") + 
  labs(x = "Year", y = "Temperature (F)", color = "Measurement Type",
       title = "Greater Portland Area Average Temperatures")
#seems to use a gradient color scale because year is a date type as opposed to a factor or char
#when using char data it goes back to the rainbow color for groups

avg_by_year_plot
```

:::

::::